---
title: "ST2137 Tutorial 2"
author: "Tay Yong Qiang"
date: "31 August 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
```

## Data sets

There are five variables in the `tut2htwtfixed.txt` dataset. They are

- `id`: Identity of the subject (Columns 1-3)
- `gender`: Gender of the subject (Column 4)
- `height`: Height of the subject in cm (Columns 5-7)
- `weight`: Weight of the subject in kg (Columns 8-9)
- `siblings`: Number of siblings of the subject (Column 10)

`tut2test.csv` is an Excel file with comma-separated values. There are two variables in it. They are

- `id`: Identity of the subject
- `test`: Test score of the subject

##My solutions

1. Create an R data frame `htwt2` by importing the `tut2htwtfixed.txt` file into the R.

```{r echo = c(-1,-4)}
setwd("C:/Users/Tay/Downloads/NUS/ST/ST2137/Tutorials/Tutorial 2/Data set")
varnames = c("id", "gender", "height", "weight", "siblings")
htwt2 <- read.fwf("tut2htwtfixed.txt", widths = c(3,1,3,2,1), header = F, col.names = varnames)
```
For an example, here are the first 5 rows of `htwt2`:
```{r}
htwt2[1:5,]
```

2. Based on `htwt2`, create an R data frame `htwt2m` which contains the data for all the male subjects. How many males are there in the data frame `htwt2`?

```{r message = F}
attach(htwt2)
htwt2m <- htwt2[gender == "M",]
dim(htwt2m)
```
There are 48 males in `htwt2`.
Note that it is important to `attach()` the data to make the variables accessible by name. Without this step, it would not make sense to call 'gender' in the next command.

3. Import `tut2test.csv` into the R. Merge the two datasets `htwt2` and `tut2test`. Let us call this new R data frame `htwttest2`. Identify individuals whose height is greater than 182 cm. What are the test scores of subjects whose height is greater than 182 cm?

```{r echo = c(-1), message = F}
setwd("C:/Users/Tay/Downloads/NUS/ST/ST2137/Tutorials/Tutorial 2/Data set")
tut2test <- read.csv("tut2test.csv", header = T)
htwttest2 <- merge(htwt2, tut2test, by = "id")
attach(htwttest2)
higherThan182 <- htwttest2[height > 182,]
higherThan182
```

4. Suppose that there was an error in the weight of the Subject 210 in the text file. Obtain a new R data frame `htwttest2remo` by removing the record related to the Subject 210 from the data frame `htwttest2`.

```{r echo = c(-1,-3)}
start <- htwttest2[1,1]
htwttest2remo <- htwttest2[id != 210,]
htwttest2remo[(210-2-start):(210+2-start),]
```
Subject 210 has been removed!

5. After checking with the Subject 210, we found out that his actual weight should be 68 kg instead of 88 kg. Modify the R data frame ``htwttest2` by rectifying the mistake.

```{r echo = c(-3)}
htwttest2remo <- htwttest2[id != 210,]
htwttest2[210 - htwttest2[1,1] + 1, 4] <- 68
htwttest2[(210-start+1),]
```
The weight has been changed.

6. Who is the second tallest female in this group? What are her height, weight, and test score?

```{r message = F}
htwttest2female <- htwttest2[gender == "F",]
attach(htwttest2female)
htwttest2female[rev(order(height)),][2, c("height", "weight", "test")]
```

7. Create a new variable called ``grade` using the following rules: (1) grade = “A” if test ≥ 80, (2) grade = “B” if 70 ≤ test < 80, (3) grade = “C” if 60 ≤ test < 70, (4) grade = “D” if 50 ≤ test < 60 and (5) grade = “F” if test < 50. How many subjects who have “F” grade are there?

```{r message = F}
htwttest2["grade"] <- ""
attach(htwttest2)
for (i in 1:nrow(htwttest2)){
  score = htwttest2[i,"test"]
  if (score >= 80){
    htwttest2[i, "grade"] = "A"
  }
  else if (score >= 70){
    htwttest2[i, "grade"] = "B"
  }
  else if (score >= 60){
    htwttest2[i, "grade"] = "C"
  }
  else if (score >= 50){
    htwttest2[i, "grade"] = "D"
  }
  else htwttest2[i, "grade"] = "F"
}
```

Here are the 1st 5 rows of the dataset after grade assignment:

```{r}
htwttest2[1:5,]
```

8. Suppose a matrix $X=\left(\begin{array}{cc}{1} & {1} \\ {1} & {3} \\ {1} & {4} \\ {1} & {7} \\ {1} & {11}\end{array}\right)$ and $\underline{y}=\left(\begin{array}{c}{4} \\ {6} \\ {13} \\ {15} \\ {19}\end{array}\right)$.
Define $\hat{\underline{\beta}}=\left(X^{\prime} X\right)^{-1} X^{\prime} \underline{y} .$ Using the matrix operations in $\mathrm{R}$ to find $\hat{\underline{\beta}}$.

```{r}
X <- cbind(c(1,1,1,1,1), c(1,3,4,7,11))
y <- c(4,6,13,15,19)
beta <- solve(t(X) %*% X) %*% t(X) %*% y
X; y; beta
```

9. A sequence is generated using the following recursive relation
$$x_{n}=x_{n-1}-2 x_{n-2} \quad \text{ for } n \geq 3,\ x_{1}=0 \text { and } x_{2}=2.$$
(i) Use the loop function in R to find the 18th term of the series.
(ii) Find the sum of the first 15 terms in this sequence.

```{r}
x_n_2 = 0
x_n_1 = 2
sum = x_n_2 + x_n_1

for (i in 3:18){
  x_n = x_n_1 - 2*x_n_2
  
  if (i == 18){
    x_18 = x_n
  }
  if (i<=15){
    sum = sum + x_n
  }
    
  x_n_2 = x_n_1
  x_n_1 = x_n
}
cat("The 18th term is", x_18, ". The sum of the first 15 terms in this sequence is", sum, ".")
```

10. Write a function that will calculate the mean, the second, the third and the fourth central moments a given data vector on variable $X .$
The $r^{\text {th }}$ central moment for $r \geq 2$ is given by
$$M_{r}=\frac{1}{n} \sum_{i=1}^{n}\left(X_{i}-\overline{X}\right)^{r}, \text { where } \overline{X}=\frac{1}{n} \sum_{i=1}^{n} X_{i}$$
Hence obtain the mean, the second, the third and the fourth central moments for the height in the data frame `htwt2`.

```{r}
mean <- function(x){
  n <- length(x)
  return(sum(x) / n)
}
rth_moment <- function(r, x){
  n <- length(x)
  moment <- sum((x - mean(x)) ^ r) / n
  return(moment)
}
calc_moments <- function(x){
  stats <- c(mean(x), rth_moment(2,x), rth_moment(3,x), rth_moment(4,x))
  return(stats)
}

cat("The first 4 central moments are", calc_moments(htwt2$height), "respectively.")
```